hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-03-03 inst/manuscript/paper.Rmd
predicting phenology edits 2

^2^Idaho Fish and Wildlife Conservation Office, U.S. Fish and Wildlife Service, Orofino, ID, USA

@@

```{r, echo = TRUE}
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  development_type = "hatch"
  )
```

@@

```{r, warning=FALSE, echo = TRUE}
WI_hatch <- predict_phenology(
  data = woody_island,
  dates = date,
  temperature = temp_c,
  spawn.date = "1990-08-18",
  model = sockeye_hatch_mod
  )
```

@@

```{r, eval=FALSE, echo = TRUE}
str(WI_hatch)
```

@@

`WI_hatch$model_specs` -- Provides details about the model used for prediction, including whether it was retrieved from `model_select()` or generated using `fit_model()`. Most importantly, it contains the model expression (i.e., the formula) used for phenology predictions.

>>>>

^2^Clearwater Solutions, Lenore, ID, USA

@@

```{r, echo = TRUE}
#model_select() to get sockeye model
sockeye_hatch_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  development_type = "hatch"
  )
```

@@

```{r, warning=FALSE, echo = TRUE}
#predict_phenology() with sockeye model and woody_island temperature data
WI_hatch <- predict_phenology(
  data = woody_island,
  dates = date,
  temperature = temp_c,
  spawn.date = "1990-08-18", #notice the character string for spawn date
  model = sockeye_hatch_mod
  )
```

@@

```{r, echo = TRUE}
summary(WI_hatch)
```

@@

`WI_hatch$model_specs` -- Provides details about the model used for prediction, including whether it was retrieved from `model_select()` or generated using `fit_model()`. Most importantly, it contains the model expression (*i.e.*, the formula) used for phenology predictions.
